---
output:
  md_document:
    variant: gfm
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

**This syllabus is effective as of `r strftime(Sys.time(), "%A,  %B %d, %Y at %I:%M %p", tz = "America/New_York")`**

# DNSC 6278 Big Data Analytics<br/>the George Washington University<br/>Summer 2019

## Course Information

- **Instructors:** Marck Vaisman (marck at gwu.edu)
- **Classroom:** Duques 255
- **Time:** Wednesday 6:00-9:00pm
- **TA's:** Swetha Kalla (swetha at gwmail.gwu.edu)
- **TA Office Hours:** TBD
<br/>
- **Course Description:** Data is everywhere! Many times, it's just too big to work with traditional tools. This is a hands-on, practical workshop style course about using cloud computing resources to do analysis and manipulation of datasets that are too large to fit on a single machine and/or analyzed with traditional tools. The course will focus on Spark, MapReduce, the Hadoop Ecosystem and more.<br/>
You will understand how to acquire and/or ingest the data, and then massage, clean, transform, analyze, and model it within the context of big data analytics. You will be able to think more programmatically and logically about your big data needs, tools and issues. 
- **Credit Hours:** This is a 3 credit graduate level course. You will spend approximately 3 hours per week in class. It is expected that you will spend approximately 2-3 hours of outside classroom activities (required readings, homework problems, completion of labs) for each hour of class time. You will spend approximately 30 hours in instructional time, and approximately 100 hours in out-of-classroom time.
- **Learning Objectives:**
  - Operate big data tools and cloud infrastructure, including Spark, MapReduce, Hadoop and other tools in the big data ecosystem 
  - Recognize and use ancillary tools that support big data processing, including git and the Linux command line
  - Setup and manage big data infrastructure and tools in the cloud on Amazon Web Services and Microsoft Azure
  - Identify broad spectrum resources and documentation to remain current with big data tools and developments
  - Execute a big data analytics exercise from start to finish: ingest, wrangle, clean, analyze and store
  - Be aware of the responsibilities that are associated with performing analysis of large datasets 

### Pre-requisites

- Experience with Python and SQL.  **Note:** We will use Python as the primary interface to Apache Spark, through [PySpark](https://spark.apache.org/docs/latest/api/python/index.html)
- Understand programming concepts (flow control, input/output, variable assignment.) 
- Experience with git and GitHub

### Refresher Tutorials

It is highly recommended that you go through the following tutorials if you need a refresher or are new to the topics of git, the command line, and SQL. If these topics are new to you, you must take these online courses before the course begins.

- [git - the simple guide](http://rogerdudler.github.io/git-guide/)
- [codeacademy - Learn the Command Line](https://www.codecademy.com/learn/learn-the-command-line)
- [codeacademy - Learn SQL](https://www.codecademy.com/learn/learn-sql)

## Books, Software and Cloud Resources



We will also provide supplemental materials (articles, links, videos, etc.) to complement the books. 

### Books (for assigned readings)

We have chosen several reference books for this course that cover different parts of the material. We will assign readings for each class from these books. These books are available on [Safari Books Online](https://www.safaribooksonline.com/), and you should be able to access these resources. Our understanding is that as a GWU student, you have access to these resources. 

- Benjamin Bengfort, Jenny Kim (2016). _Data Analytics with Hadoop: An Introduction for Data Scientists_. O'Reilly Media. ISBN: 9781491913703. 
- Bill Chambers, Matei Zaharia (2018). _Spark: The Definitive Guide_. O'Reilly Media. ISBN: 9781491912218.

### Papers

We have chosen several seminal papers in the big data space to complement the book. The papers discuss the tools and history in much more detail. Topics in the papers will be included in the final exam. 

### Cloud Resources

You will be using cloud resources on [Microsoft Azure](https://azure.microsoft.com/en-us/) and [Amazon Web Services](http://aws.amazon.com). We will discuss how to setup your account and environment in the first class session.

### Terminal and Command Line

You will be using the command line and terminal heavily in this course. Please take a look at the following depending on your laptop's operating system

- **For Windows users:** Read [this blog post](https://www.howtogeek.com/336775/how-to-enable-and-use-windows-10s-built-in-ssh-commands/) to enable [ssh](https://en.wikipedia.org/wiki/Secure_Shell) commands on Windows10. Even if you have PuTTY or anothe ssh client installed, you will be using the command line.
- **For Mac users:** Macs have a built in Terminal. However, I prefer using [iTerm](https://www.iterm2.com/) as another Terminal application for your Mac. I've been using it for years and I love it. This is not required, but truly recommended.
- **Linux users:** nothing to do. 

### Credit Cards

You will need a credit card or a debit card to create an account on Amazon Web Services. If you do not have a credit card, you may consider getting a pre-paid VISA card which you can use as the credit card when you create the account.


## Learning Activities, Communication and Evaluation

This is a hands-on, practical, workshop style course that provides opportunities to use the tools and techniques discussed in class. Although this is not a programming course per se, there is programming involved. 

### Lectures and In-Class Labs

Every class session will have a lecture portion and mosy sessions will have an in-class lab portion. Lab exercises are designed to get you familiar with the tools discussed in class. In these labs, we will work through simple examples. The completion of the in-class lab exercises is part of your attendance/participation portion of the grade. 

### Piazza

This term we will be using Piazza for class discussion. The system is highly catered to getting you help fast and efficiently from classmates, the TA, and myself. Rather than emailing questions to the teaching staff, I encourage you to post your questions on Piazza. 

Find our class page at: [https://piazza.com/gwu/summer2019/dnsc6278/home](https://piazza.com/gwu/summer2019/dnsc6278/home)

### Assignments

You will be given problem sets as homework assignments. The goal of these problem sets is to use the big data tools to answer some questions about large datasets. The problem sets will build on the labs and will be much more elaborate. Deliverables from the problem sets will usually include code written for your programs and the output produced.

**Important Note:** We reuse problem set questions, we expect students not to copy, refer to, or look at the solutions in preparing their answers. Since this is a graduate-level class, we expect students to want to learn and not search online for answers.

### Group Project 

Students will assemble into groups of 2-4 students and propose, perform, and write up an analysis of a _big_ dataset using the tools learned in class. _Big_ is defined as "a dataset that is so large that you cannot work with it on a laptop."

Details for the project can be found [here](project.md).

### Final Exam

There will be an in-class final exam during the last class session. It will be closed everything (no books, slides, computers, etc.) You will only be allowed to bring along with you 1 letter-sizes paper of your own notes ('cheat sheet'), you can use both sides. The exam will be comprehensive and will cover topics from the readings and in-class discussion. 

### Grading

- Problem Sets: 50 points (5 individual assignments at 10 points each)
- Group Project: 30 points
- Final Exam: 20 points

Final point count is 100. We have no plans to curve the final grade, so the letter grade will follow standard guidelines:

- A: >= 93
- A-: >= 90, < 93
- B+: >= 87, < 90
- B: >= 83, < 87
- B-: >= 80, < 83
- C: < 80

## Course Calendar 

This calendar is subject to change. We will make make any changes known in advance.

In the readings, "DAWH" is _Data Analytics with Hadoop_, and "Spark" is _Spark: The Definitive Guide_. Papers are referenced by author and title.

```{r, echo = FALSE, results="asis"} 
source("schedule.R")
sched %>%
  knitr:::kable(., "markdown",
                col.names = c("Class", "Date", "Topics", "Lab",
                              "Readings", "Notes")
  )

``` 


## Policies & Expectations

### General Policies

- **Attendance and punctuality:** Given the technical nature of this course, and the breadth of topics discussed, you are expected to attend each class, to complete all readings, and to participate actively in lectures, discussions and exercises. Please inform us in advance if you are not able to attend class for any reason.
- **Participation:** We love participation. Read. Raise your hand. Ask questions. Make comments. Challenge us. Acknowledge us. If we speak for three hours to a silent classroom, it is a lot more boring and tiring for everyone.
- **Computer Usage:** You must bring your laptop to class to work on labs. However, please limit the use of computers and phones to the lab portion of the class and please refrain from other activities. **You may not use your computer or phone while your peers or guest speakers are presenting.**
- **Email and Online Discussion Boards:** Please use the [discussion boards on Piazza](https://piazza.com/gwu/summer2019/dnsc6278/home) for questions about the course, homework assignments, technical issues, etc. Staff will be monitoring them and providing answers on a regular basis. Individual emails will not be answered except for special circumstances.
- **Cloud Resources:** You will create cloud accounts on Amazon Web Services and Microsoft Azure. You will get credits on both platforms that will be enough to support your coursework throughout the semester. **It is your responsibility to manage the credits and resources yourself. If you run out of credits because you do not shut down your resources, we cannot help you.**
- **Allocate Time for Assignments:** Homeworks take time, so please do not wait until the last minute to start them. Give yourself several days to work on problem sets. While the tools have matured a lot over the years, there are cases where you will run into unforseen technical difficulties. All homework assignments have been thoroughly tested using the technical configuration provided in the assignment and they work. _"It didn't work for me"_ is not an excuse. _"I lost my code because I didn't push to github"_ is not an excuse. _"It took me too long because it was the first time I'm doing it"_ is not an excuse. 
- **Late Policy:** Assigments are due before class at 6pm on the due date. Due dates will not be extended, and late homeworks will incur a late penalty of 25% per day. 
- **Class materials are for class use only!:** Please refrain from making your private GitHub repositories or any other class materials public. 

### Open Door Policy

Please approach or get in touch with us if something is not working for you regarding the class, methods, etc. Our pledge to you is to provide the best learning experience possible. If any issue comes up, please do not wait until the last minute to bring it up. We will work with you individually to try to resolve in a timely manner.

## Academic Integrity

The code of academic integrity applies to all courses in the George Washington School of Business. Please become familiar with the code. All students are expected to maintain the highest level of academic integrity throughout the course of the semester. Please note that acts of academic dishonesty during the course will be prosecuted and harsh penalties may be sought for such acts. Students are responsible for knowing what acts constitute academic dishonesty. The code may be found at [http://www.gwu.edu/~ntegrity/code.html](http://www.gwu.edu/~ntegrity/code.html)

### Collaboration Policy

* **In-class labs:** you may collaborate
You may collaborate with other students during in-class labs to facilitate collective learning. 
* **Group project:** by nature, it is a group project and collaboration is to be confined within groups. **You may not collaborate across groups.**
* **Individual assignments:**
  * You may discuss ideas on how to solve the problem with other students
  * You may get help from other students if you are stuck and/or are having technical difficulties on steps to take
  * **You may not share code - this will be checked**
  * **All submitted code and scripts written must be your own - this will be checked**
  * **You must do all your work on your own cloud resources - this will be checked**

**We have a ZERO TOLERANCE POLICY.**

## University Policies and Support Services 

### Religious Accommodation

Students should notify faculty during the first week of the semester of their intention to be absent from class on their day(s) of religious observance. Faculty should extend to these students the courtesy of absence without penalty on such occasions, including permission to make up examinations.  Faculty who intend to observe a religious holiday should arrange at the beginning of the semester to reschedule missed classes or to make other provisions for their course-related activities.

### Disability Support Services (DSS)

Any student who may need an accommodation based on the potential impact of a disability should contact the Disability Support Services office at 202-994-8250 in the Rome Hall, Suite 102, to establish eligibility and to coordinate reasonable accommodations. For additional information please refer to [http://gwired.gwu.edu/dss/](http://gwired.gwu.edu/dss/)

### Mental Health Services 202-994-5300

The University's Mental Health Services offers 24/7 assistance and referral to address students' personal, social, career, and study skills problems. Services for students include: crisis and emergency mental health consultations confidential assessment, counseling services (individual and small group), and referrals. [http://counselingcenter.gwu.edu](http://counselingcenter.gwu.edu)


